---
title: NYC Taxi Trip Duration Code Improvement
author: Jiao Jiang 
output: html_document
---
######## Load required packages

```{r, echo=FALSE}
installIfAbsentAndLoad <- function(neededVector) {
  for(thispackage in neededVector) {
    if( ! require(thispackage, character.only = T) )
    { install.packages(thispackage)}
    require(thispackage, character.only = T)
  }
}

```

```{r}
needed <- c("data.table", "tidyverse","lubridate","gridExtra","ggplot2","corrplot") 
```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
installIfAbsentAndLoad(needed)
```
######## Read Dataset
```{r}
train <- read.csv("train.csv")
test <- read.csv("test.csv")
count(train)
```
######## 1,458,641 data points in train dataset

######## Improvement 1: Data Cleaning
######## 1.1 Look for missing values
```{r}
sum(is.na(train))
sum(is.na(test))
```

######## Examine Train Dataset
```{r}
summary(train)
```

######## 1.2 Look for value that doesn't make sense
######## trip_duration is measured in seconds
```{r}
day_plus_trips <- train %>% filter(trip_duration >= 12*60*60)
count(day_plus_trips)
```
######## 1993 observations
```{r}
less_than_1_min_trips <- train %>%filter(trip_duration <= 60)
count(less_than_1_min_trips)
```
######## 8777 observations

######## get rid of day_plus_trips & less_than_1_min_trips
######## reduce size by 0.7%
```{r}
train=train[!row.names(train) %in% row.names(day_plus_trips),]
train=train[!row.names(train) %in% row.names(less_than_1_min_trips),]
```

######## 1.3 Fix Date Variables
```{r, warning=FALSE, error=FALSE}
train = as.data.table(train)
train[,pickup_datetime:=as.Date(pickup_datetime)]
train[,dropoff_datetime:=as.Date(dropoff_datetime)]
train[,":="(
  pickup_yday=yday(pickup_datetime)
  ,pickup_mday=mday(pickup_datetime)
)]
train$pickup_week <- week(train$pickup_datetime)
train$pickup_month <- month(train$pickup_datetime)
train$pickup_weekdays <- weekdays(train$pickup_datetime)
train$pickup_weekend <- ifelse(train$pickup_weekdays==1 | train$pickup_weekdays==7,"Weekend","not-Weekend")
```
######## Plot
######## 1.3.1 Mean trip duration by month  
```{r}
plot1 = train[, list(mean_trip_duration= mean(trip_duration)), by=pickup_datetime] %>%
  ggplot(aes(x=pickup_datetime, y=mean_trip_duration)) + 
  geom_bar(stat='identity', fill='steelblue') + 
  labs(x='', y='Mean Trip Duration', title='Mean Trip Duration over time')
grid.arrange(plot1)
```

######## 1.3.2 Mean trip duration by weekdays 
```{r}
plot2 <-train[, list(mean_trip_duration = mean(trip_duration)), by = pickup_weekdays] %>%
  ggplot(aes(x = pickup_weekdays, y = mean_trip_duration)) +
  geom_bar(stat = 'identity', fill = 'steelblue') +
  labs(x = 'Month', y = 'Mean Trip Duration', title = 'Mean Trip duration by weekdays')
grid.arrange(plot2)
```

######## 1.4 correlation matrix
```{r}
corr_features = train[,.(vendor_id, pickup_week, pickup_month,pickup_yday, pickup_mday,
                         passenger_count,pickup_longitude, pickup_latitude, dropoff_longitude,
                         dropoff_latitude, trip_duration)] 
corrplot(cor(corr_features, use='complete.obs'), type='lower')
```




